### 2022-03-01 paper/include/structure.tex

\section{Structure}
After copying the template to your git repository make sure you also
copied the (hidden) \texttt{.gitignore} file.
You also find additional directories which are explained in the following:

\begin{figure*}
		\centering
	\begin{subfigure}[c]{0.45\textwidth}
				\centering
		\includegraphics[width=.4\textwidth]{pics/example}
		\subcaption{An external figure.}
		\label{fig:external_pdf}	
	\end{subfigure}
	\begin{subfigure}[c]{0.45\textwidth}
		\includegraphics[width=\textwidth]{figures/example}
		\subcaption{An external figure which was built before included.}
		\label{fig:external_tex}
	\end{subfigure}
	\caption{This Figure has two subfigures and spans both columns.}
\end{figure*}

\begin{itemize}
	\item[\texttt{/pics}:] Here you can put all external figures which are
	not built for the paper. These are, for example, PNG or JPG files and
	vector graphics like PDF files. For an example have a look at
	Figure~\ref{fig:external_pdf}.

	\item[\texttt{/figures}:] If you do not want to build all your figures
	every time you compile your paper it is helpful to build them
	separately. All \texttt{.tex} and \texttt{.tikz} files within this directory
	are built automatically by the Makefile. Since these figures are built
	independently of the main document they have to be of type
	\texttt{standalone}. Make sure that you put no \texttt{.tex} or
	\texttt{.tikz} inside this directory which cannot be compiled
	independently.
	Figure~\ref{fig:external_tex} is an example for this. Notice that this
	PDF is only built if the figures source file has changed.

	\item[\texttt{/data}:] If you have larger amounts of data which you
	want to put into your paper, for example in form of a table or a plot
	you can import the data directly from another file.
	Figure~\ref{fig:external_tex} reads the red sample points from a CSV
	file located in this directory.
\end{itemize}