### 2022-03-01 paper/paper.tex

% Do not change the section below
% ============================================================================
\documentclass[conference,compsoc,a4paper]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{nimbusmononarrow}
\usepackage[T1]{fontenc}

\usepackage{balance}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{url}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\newcommand{\fixme}[1]{\textcolor{red}{[FIXME: #1]}}
% ============================================================================

% Add additionale LaTeX packages if needed here
% ============================================================================
\usepackage[%
	detect-all,
	binary-units=true,
]{siunitx}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{moeptikz}
% ============================================================================

\begin{document}

% Add paper title here, most likely you want to delete the \thanks{} macro
% ============================================================================
\title{Paper Title}
% ============================================================================

% Update author information
% ============================================================================
\author{%
	\IEEEauthorblockN{%
		Your Name,
		Advisor's Name\IEEEauthorrefmark{1}
	}
	\IEEEauthorblockA{%
		\IEEEauthorrefmark{1}Chair of Network Architectures and Services,
		Department of Informatics\\
		Technical University of Munich, Germany\\
		Email: your@mail.address, advisor@net.in.tum.de
	}
}
% ============================================================================

\maketitle

% Edit include/abstract.tex to write your paper's abstract
% ============================================================================
\begin{abstract}
\input{include/abstract}
\end{abstract}
% ============================================================================

% List keywords of your paper in lower-case below
% ============================================================================
\begin{IEEEkeywords}
software-defined networks, measurement, hight-speed networks
\end{IEEEkeywords}
% ============================================================================

% For each section of your chapter add one source file in the include folder
% ============================================================================
\input{include/introduction}
\input{include/related_work}
\input{include/formatting}
\input{include/structure}
\input{include/conclusion}
% ============================================================================

% Put your references in the paper.bib file
% ============================================================================
\balance

\bibliographystyle{IEEEtran}
\bibliography{paper}
% ============================================================================

\end{document}