### 2022-11-04 paper/include/structure.tex
remove necessity for make

\section{Structure}
After copying the template to your git repository make sure you also
copied the (hidden) \texttt{.gitignore} file.
You also find additional directories which are explained in the following:

\begin{figure*}
		\centering
	\begin{subfigure}[c]{0.45\textwidth}
				\centering
		\includegraphics[width=.4\textwidth]{pics/example}
		\subcaption{An external figure.}
		\label{fig:external_pdf}	
	\end{subfigure}
\end{figure*}

\begin{itemize}
	\item[\texttt{/pics}:] Here you can put all external figures which are
	not built for the paper. These are, for example, PNG or JPG files and
	vector graphics like PDF files. For an example have a look at
	Figure~\ref{fig:external_pdf}.

	\item[\texttt{/data}:] If you have larger amounts of data which you
	want to put into your paper, for example in form of a table or a plot
	you can import the data directly from another file.
\end{itemize}