### 2022-11-04 talk/include/slides.tex

\usepackage[utf8]{inputenc}
\usepackage{tumcd/packages}
\usepackage{tumcd/beamermods}

% For beamer mode (default):
\author[S. G\"unther]{Stephan Günther}
\title[Networking]{Something with Networking}

% Uncomment to add advisors for presentation in Oberseminar
%\advisor{Herrmann B. Wolfgang}

% Uncomment to add type for presentation in Oberseminar
% Usage: \thesistype{intermediate | final}{bachelor | master | idp | gr}
%\thesistype{intermediate}{gr}

% Uncomment to configure date in dd-mm-yyyy format
%\newdate{date}{25}{05}{2017}
%\date{\displaydate{date}}

% Uncomment to add a venue to the title slide
%\venue{International Conference on Conferencing 2018 \\ Garching b. München, Germany}

% For lecture mode (use package option 'lecture'):
%\lecture[GRNVS]{Grundlagen Rechnernetze und Verteilte Systeme}
%\module{IN0010}
%\semester{SoSe\,2016}
%\assistants{Johannes Naab, Stephan Günther, Maurice Leclaire}

\usepackage{pgfpages}
\usepackage{ifthen}
% ============================================================================
% jobname solution
% ============================================================================
\newif\ifsolution%
\ifthenelse{\equal{\detokenize{notes}}{\jobname}}{%
\setbeameroption{show notes on second screen=bottom}
\setbeamercolor{note page}{bg=white, fg=black}
\setbeamercolor{note title}{bg=white!95!black, fg=black}
}{
}

% TeXLive 2018 compatibility: https://tex.stackexchange.com/questions/426088/texlive-pretest-2018-beamer-and-subfig-collide
\makeatletter
\let\@@magyar@captionfix\relax
\makeatother