### 2024-11-20 paper/figures/mitm.tex

\begin{tikzpicture}[
    scale=0.4, % Further reduced scale
    node distance=2cm, % Further reduced distance
    user/.style={circle,draw,minimum size=0.6cm},
    attacker/.style={rectangle,draw,fill=red!20,minimum width=1cm,minimum height=0.7cm},
    server/.style={rectangle,draw,minimum width=1cm,minimum height=0.7cm}
]
    % Define nodes with shorter labels
    \node[user] (alice) {A};
    \node[attacker] (mitm) [right of=alice] {M};
    \node[server] (server) [right of=mitm] {S};

    % Shorter arrows and smaller labels
    \draw[->,thick] (alice) to[bend left=20] node[above,font=\tiny] {Msg} (mitm);
    \draw[->,thick] (mitm) to[bend left=20] node[above,font=\tiny] {Mod} (server);
    \draw[->,thick] (server) to[bend left=20] node[below,font=\tiny] {Rsp} (mitm);
    \draw[->,thick] (mitm) to[bend left=20] node[below,font=\tiny] {Mod} (alice);
\end{tikzpicture}